---
title: "case_study_2"
author: "sunna quazi and patrick mcdevitt"
date: "21 avril 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## __Case Study 2__  







***  
#### __Question 03 (20 points)__  

The built-in data set called Orange in R is about the growth of orange trees. The Orange data
frame has 3 columns of records of the growth of orange trees.  
Variable description  
* Tree : an ordered factor indicating the tree on which the measurement is made. The ordering
is according to increasing maximum diameter.  
* age : a numeric vector giving the age of the tree (days since 1968/12/31)  
* circumference : a numeric vector of trunk circumferences (mm). This is probably
“circumference at breast height”, a standard measurement in forestry.  
	a) Calculate the mean and the median of the trunk circumferences for different size of the
trees. (Tree)  
	b) Make a scatter plot of the trunk circumferences against the age of the tree. Use different
plotting symbols for different size of trees.  
	c) Display the trunk circumferences on a comparative boxplot against tree. Be sure you order the boxplots in the increasing order of maximum diameter.  
	
Submit your final R code and necessary plots for each part.


```{r}
library(ggplot2)
data (Orange)

plot(circumference ~ age, data = Orange,
     xlab = "Tree age (days since 1968/12/31)",
     ylab = "Tree circumference (mm)", las = 1,
     main = "Orange tree data and fitted model",
	 col = Tree)


p <- ggplot(Orange, aes(age, circumference))
p + geom_point(aes(color = factor(Tree)))

Orange$tree <- as.numeric(Orange$Tree)

p <- ggplot(Orange, aes(x = as.numeric(Tree), y = circumference))
p + geom_boxplot(aes(color = factor(tree)))

```





***  
##### __... Question 05 (10 points)__  
##### ... Semester End Bonus  
##### ... Run the following code in R and paste the final graph.  


```{r q5, echo=FALSE}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	t<-seq(0,10,length=1000)
	x<-sqrt(t)*cos(2*pi*t)
	y<-sqrt(t)*sin(2*pi*t)
	
	plot(x,y,axes=F,type="l",lwd=3,xlab="x(t)",ylab="y(t)",col="red")
	axis(1,at=seq(-3,3,by=0.5),labels=seq(-3,3,by=0.5))
	axis(2)
	
	box()
	
	title(main=expression(
	paste("(x(t),y(t)) with polar coordinates",
	(list(sqrt(t),2*pi*t))
	)))
```