---
title: "Case Study 2, Q4"
author: "squazi and Patrick"
date: "April 25 2017"
output:
  html_document:
    keep_md: true
---
#Temp Data
## I
```{r}
#load temp data
temp <- read.csv("C:/Users/esunqua/Documents/R/Case Study 2/TEMP.csv")
str(temp)
head(temp)
#Observation- data is monthly temperatures for each country for 3 to 4 centuries
```
###Find the difference between thde maximum and the minimum monthly average temperatures for each country and report/visualize top 20 countries with the maximum differences for the period since 1900.

####We need to get first get rid of all the pre-1900 values by subsetting
```{r}
#Observation: dates are in all sorts of different formats, but the dates that I want (the 1900+ dates) have "/" instead of "-"
temp$Date2 <-as.character(temp$Date)
substring="/"
tempSub <-subset(temp, grepl(substring,temp$Date2))
tempSub$Date3 <-as.Date(tempSub$Date, format = "%m/%d/%Y")
head(tempSub)
```
####Now we need to get the maximum and minimum monthly values for each country.
```{r}
#Get rid of NA values, because they won't contribute to the min or max
library(plyr)
tempClean <-subset(x=tempSub,  !is.na(Monthly.AverageTemp))
#Get the mins and max per country
x <-ddply(tempClean, c("tempClean$Country"), function(x) summary(x$Monthly.AverageTemp))
#Create a new variable that has teh difference between the max an min temperatures per country
x["Difference"] <- x$Max.-x$Min.
#Sort the new difference variable in ascending order
xSort <-arrange(x,desc(Difference))
```
###Report the Top 20 countries with the largest range in monthly temperatures:
```{r}
xSort$Country <-xSort$'tempClean$Country'
Ifinal <-print(xSort[("1":"20"), c("Country", "Max.", "Min.", "Difference")])
```
###Visualize the Top 20 countries with the largest range in monthly temperatures:
```{r}
Ifinal$Country <-as.character(Ifinal$Country)
country <-Ifinal$Country
Difference <-Ifinal$Difference
counts <-table(rep(country, Difference))
barplot(counts, main="Top 20 Countries with the Highest Temperature Difference",col = rainbow(20), ylim=c(0,50), ylab="Diff of Monthly Avg. Temps.", las=2)
```
##II
###Select a subset of data called "UStemp" where US land temperatures from 01/01/1990 in Temp data. Use UStemp dataset to answer the followings.
```{r}
tempClean$Country <-as.character(tempClean$Country)
UStemp <-subset(tempClean,subset = (Country == "United States"))
```
###a) Create a new column to display the monthly average land temperatures in Fahrenheit (°F).
```{r}
UStemp$Fahrenheit <- (UStemp$Monthly.AverageTemp*1.8)+32
```
###b) Calculate average land temperature by year and plot it. The original file has the average land temperature by month.
```{r}
#tidy up UStemp
UStemp$Date <-UStemp$Date3
UStemp <-subset(UStemp, select = -Date2)
UStemp <-subset(UStemp, select = -Date3)
UStemp$Year<-format(UStemp$Date,'%Y')
ii <-ddply(UStemp, c("Year"), function(x) mean(x$Fahrenheit))
```
####Scatter Plot Method
```{r}
year = ii$Year       
average = ii$V1         
plot(year, average, main ="Modern Yearly Average Temperature in the US", 
   xlab="Year",       
   ylab="Average Yearly Temperature", pch = 20, col=rev(heat.colors(114)))
```
####Boxplot
```{r}
UStemp$Y2 <-as.numeric(UStemp$Year)
UStemp$Decade <-(UStemp$Y2 - (UStemp$Y2 %% 10))
boxplot(UStemp$Fahrenheit~UStemp$Decade, las = 2, xlab="Decade", main ="Average Temperature in the US", col=rev(heat.colors(12)))
```
###c) Calculate the one year difference of average land temperature by year and provide the maximum difference (value) with corresponding years.
```{r}
change <-diff(ii$V1)
change <-c(0,change)
ii$Change <-change
summary(ii$Change)
```
##iii
### Download "CityTemp" data set (check your SMU email). Find the difference between the maximum and the minimum temperatures for each major city and report/visualize top 20 cities with maximum differences for the period since 1900.
```{r}
citytemp <-read.csv("C:/Users/esunqua/Documents/R/Case Study 2/CityTemp.csv")
str(citytemp)
#Dates are factors again
citytemp$D2 <-as.character(citytemp$Date)
CitySub <-subset(citytemp, grepl(substring,citytemp$D2))
CitySumm <-ddply(CitySub, c("CitySub$City"), function(x) summary(CitySub$Monthly.AverageTemp))
CityClean <-subset(x=CitySub,  !is.na(Monthly.AverageTemp))
CitySumm <-ddply(CityClean, c("CityClean$City"), function(x) summary(x$Monthly.AverageTemp))
CitySumm$diff <-CitySumm$Max.-CitySumm$Min.
CitySumm <-arrange(CitySumm,desc(diff))
CitySumm$City <-CitySumm$`CityClean$City`
iii <-print(CitySumm[("1":"20"), c("City", "Max.", "Min.", "diff")])
cit <-as.character(iii$City)
dif <-iii$diff
counts2 <-table(rep(cit, dif))
barplot(counts2, main="Top 20 Cities with the Highest Temperature Difference",col = rainbow(20), ylim=c(0,55), ylab="Diff of Monthly Avg. Temps.", las=2)
```
###(iv) Compare the two graphs in (i) and (iii) and comment it.
```{r}
par (mfrow=c(1,2))
barplot(counts, main="Countries",col = rainbow(20), ylim=c(0,60), ylab="Diff of Monthly Avg. Temps.", las=2)
barplot(counts2, main="Cities",col = rainbow(20), ylim=c(0,60), ylab="Diff of Monthly Avg. Temps.", las=2)
#this is an interesting comment
```