---
title: "case_study_2"
author: "sunna quazi and patrick mcdevitt"
date: "21 avril 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)

home_dir <- "~/_smu/_src/case_study_2"

data_dir <- paste0(home_dir, "/data")
setwd(home_dir)

```

## __Case Study 2__  


***  
### __Question 01__  

Create the X matrix and print it from SAS, R, and Python.

		[4 5 12]
	X = [1 0 35]
		[2 1 82]

```{r}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	in R, 1st as a data frame and 2nd as a matrix
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	x1 <- c(4, 5, 12)
	x2 <- c(1, 0 , 35)
	x3 <- c(3, 1, 82)
	df_X <- rbind.data.frame(x1, x2, x3)
	names(df_X) <- c("c1", "c2", "c3")
	df_X

	A = matrix( 
		c(4, 5, 12, 1, 0, 35, 3, 1, 82), # the data elements 
		nrow = 3,              # number of rows 
		ncol = 3,              # number of columns 
		byrow = TRUE)        # fill matrix by rows 
	A

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	Need SAS code and image here
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	Need python code and output here
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
	
	
	
	
```




***  
### __Question 02__  
Please watch videos1 and 2 in week 11 lecture assignment. You can download the code which
used for S&P from files tab.  
Please do the following with your assigned stock.  
 Download the data.  
 Calculate log returns.  
 Calculate volatility measure.  
 Calculate volatility over entire length of series for various three different decay factors.  
 Plot the results, overlaying the volatility curves on the data, just as was done in the S&P
example.  

 Group           | Stock 
---------------- | ------------- 
Patrick & Sunna  | GWPH        

```{r}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

library(tseries)

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ..	 volatility function
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

getVol <- function(d, logrets)
{
	var = 0
	lam = 0
	varlist <- c()

	for (r in logrets)
	{
		lam = lam*(1 - 1/d) + 1
		var = (1 - 1/lam)*var + (1/lam)*r^2
		varlist <- c(varlist, var)
	}
	sqrt(varlist)
}


# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	SNPdata <- get.hist.quote('gwph', quote = "Close")
	SNPret <- log(lag(SNPdata)) - log(SNPdata)
	SNPvol <- sd(SNPret) * sqrt(250) * 100


# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	recreate figure 6.12 in the text on page 155
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	volest  <- getVol(10, SNPret)
	volest2 <- getVol(30, SNPret)
	volest3 <- getVol(100, SNPret)
	
	plot(SNPdata)
	plot(log(SNPdata))
	plot(SNPret)

	plot(volest, type = "l")
	lines(volest2, type = "l", col = "red")
	lines(volest3, type = "l", col = "blue")


```

***  
### __Question 03__  

The built-in data set called Orange in R is about the growth of orange trees. The Orange data
frame has 3 columns of records of the growth of orange trees.  
Variable description  
* Tree : an ordered factor indicating the tree on which the measurement is made. The ordering
is according to increasing maximum diameter.  
* age : a numeric vector giving the age of the tree (days since 1968/12/31)  
* circumference : a numeric vector of trunk circumferences (mm). This is probably
“circumference at breast height”, a standard measurement in forestry.  
	
Submit your final R code and necessary plots for each part.


```{r}

data (Orange)

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	a) Calculate the mean and the median of the trunk circumferences
# ...		for different size of the trees. (Tree)  
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

col_drops <- c("age", "tree")
df_orange_tmp <- Orange[ , !(names(Orange) %in% col_drops)]
aggregate(df_orange_tmp[, 2], list(df_orange_tmp$Tree), mean)
aggregate(df_orange_tmp[, 2], list(df_orange_tmp$Tree), median)

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...b) Make a scatter plot of the trunk circumferences against the age of the tree.
# ...	Use different plotting symbols for different size of trees.  
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	plot(circumference ~ age, data = Orange,
		xlab = "Tree age (days since 1968/12/31)",
		ylab = "Tree circumference (mm)", las = 1,
		main = "Orange tree data",
		col = Tree)

	p <- ggplot(Orange, aes(age, circumference))
	p + geom_point(aes(color = factor(Tree)))

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	c) Display the trunk circumferences on a comparative boxplot against tree.
# ...	Be sure you order the boxplots in the increasing order of maximum diameter.  
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	Orange$tree <- as.numeric(Orange$Tree)

	p <- ggplot(Orange, aes(x = as.numeric(Tree), y = circumference))
	p + geom_boxplot(aes(color = factor(tree)))

```

***  

### __Question 04__  
Download “Temp” data set (check your SMU email)  

(i) Find the difference between the maximum and the minimum monthly average
temperatures for each country and report/visualize top 20 countries with the
maximum differences for the period since 1900.  

(ii) Select a subset of data called “UStemp” where US land temperatures from
01/01/1990 in Temp data. Use UStemp dataset to answer the followings.  

	a) Create a new column to display the monthly average land temperatures in
		Fahrenheit ( °F) .  
	b) Calculate average land temperature by year and plot it. The original file has
		the average land temperature by month.  
	c) Calculate the one year difference of average land temperature by year and
		provide the maximum difference (value) with corresponding years.  
		(for example, year 2000: add all 12 monthly averages and divide by 12 to get
		average temperature in 2000. You can do the same thing for all the available
		years. Then you can calculate the one year difference as 1991-1990, 1992-
		1991, etc)  

(iii) Download “CityTemp” data set (check your SMU email). Find the difference
between the maximum and the minimum temperatures for each major city and
report/visualize top 20 cities with maximum differences for the period since 1900.  

(iv) Compare the two graphs in (i) and (iii) and comment it.
You can use either R or Python  

```{r}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...		Question 4 needs to be complete
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...		import data sets
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	setwd(data_dir)
	temp <- read.csv (file = "TEMP.csv", stringsAsFactors = FALSE)
	citytemp <- read.csv (file = "CityTemp.csv", stringsAsFactors = FALSE)
	setwd(home_dir)
	
```
#Temp Data
## I
```{r}
#load temp data
#temp <- read.csv("C:/Users/esunqua/Documents/R/Case Study 2/TEMP.csv")
str(temp)
head(temp)
#Observation- data is monthly temperatures for each country for 3 to 4 centuries
```
###Find the difference between thde maximum and the minimum monthly average temperatures for each country and report/visualize top 20 countries with the maximum differences for the period since 1900.

####We need to get first get rid of all the pre-1900 values by subsetting
```{r}
#Observation: dates are in all sorts of different formats, but the dates that I want (the 1900+ dates) have "/" instead of "-"
temp$Date2 <-as.character(temp$Date)
substring="/"
tempSub <-subset(temp, grepl(substring,temp$Date2))
tempSub$Date3 <-as.Date(tempSub$Date, format = "%m/%d/%Y")
head(tempSub)
```
####Now we need to get the maximum and minimum monthly values for each country.
```{r}
#Get rid of NA values, because they won't contribute to the min or max
library(plyr)
tempClean <-subset(x=tempSub,  !is.na(Monthly.AverageTemp))
#Get the mins and max per country
x <-ddply(tempClean, c("tempClean$Country"), function(x) summary(x$Monthly.AverageTemp))
#Create a new variable that has teh difference between the max an min temperatures per country
x["Difference"] <- x$Max.-x$Min.
#Sort the new difference variable in ascending order
xSort <-arrange(x,desc(Difference))
```
###Report the Top 20 countries with the largest range in monthly temperatures:
```{r}
xSort$Country <-xSort$'tempClean$Country'
Ifinal <-print(xSort[("1":"20"), c("Country", "Max.", "Min.", "Difference")])
```
###Visualize the Top 20 countries with the largest range in monthly temperatures:
```{r}
Ifinal$Country <-as.character(Ifinal$Country)
country <-Ifinal$Country
Difference <-Ifinal$Difference
counts <-table(rep(country, Difference))
barplot(counts, main="Top 20 Countries with the Highest Temperature Difference",col = rainbow(20), ylim=c(0,50), ylab="Diff of Monthly Avg. Temps.", las=2)
```
##II
###Select a subset of data called "UStemp" where US land temperatures from 01/01/1990 in Temp data. Use UStemp dataset to answer the followings.
```{r}
tempClean$Country <-as.character(tempClean$Country)
UStemp <-subset(tempClean,subset = (Country == "United States"))
```
###a) Create a new column to display the monthly average land temperatures in Fahrenheit (?F).
```{r}
UStemp$Fahrenheit <- (UStemp$Monthly.AverageTemp*1.8)+32
```
###b) Calculate average land temperature by year and plot it. The original file has the average land temperature by month.
```{r}
#tidy up UStemp
UStemp$Date <-UStemp$Date3
UStemp <-subset(UStemp, select = -Date2)
UStemp <-subset(UStemp, select = -Date3)
UStemp$Year<-format(UStemp$Date,'%Y')
ii <-ddply(UStemp, c("Year"), function(x) mean(x$Fahrenheit))
```
####Scatter Plot Method
```{r}
year = ii$Year       
average = ii$V1         
plot(year, average, main ="Modern Yearly Average Temperature in the US", 
   xlab="Year",       
   ylab="Average Yearly Temperature", pch = 20, col=rev(heat.colors(114)))
```
####Boxplot
```{r}
UStemp$Y2 <-as.numeric(UStemp$Year)
UStemp$Decade <-(UStemp$Y2 - (UStemp$Y2 %% 10))
boxplot(UStemp$Fahrenheit~UStemp$Decade, las = 2, xlab="Decade", main ="Average Temperature in the US", col=rev(heat.colors(12)))
```
###c) Calculate the one year difference of average land temperature by year and provide the maximum difference (value) with corresponding years.
```{r}

change <-diff(ii$V1)
change <-c(0,change)
ii$Change <-change
summary(ii$Change)
```
##iii
### Download "CityTemp" data set (check your SMU email). Find the difference between the maximum and the minimum temperatures for each major city and report/visualize top 20 cities with maximum differences for the period since 1900.  

```{r}

#citytemp <-read.csv("C:/Users/esunqua/Documents/R/Case Study 2/CityTemp.csv")
str(citytemp)

#Dates are factors again

citytemp$D2 <-as.character(citytemp$Date)
CitySub <-subset(citytemp, grepl(substring,citytemp$D2))
CitySumm <-ddply(CitySub, c("CitySub$City"), function(x) summary(CitySub$Monthly.AverageTemp))
CityClean <-subset(x=CitySub,  !is.na(Monthly.AverageTemp))
CitySumm <-ddply(CityClean, c("CityClean$City"), function(x) summary(x$Monthly.AverageTemp))
CitySumm$diff <-CitySumm$Max.-CitySumm$Min.
CitySumm <-arrange(CitySumm,desc(diff))
CitySumm$City <-CitySumm$`CityClean$City`
iii <-print(CitySumm[("1":"20"), c("City", "Max.", "Min.", "diff")])
cit <-as.character(iii$City)
dif <-iii$diff
counts2 <-table(rep(cit, dif))
barplot(counts2, main="Top 20 Cities with the Highest Temperature Difference",col = rainbow(20), ylim=c(0,55), ylab="Diff of Monthly Avg. Temps.", las=2)

```

###(iv) Compare the two graphs in (i) and (iii) and comment it.  

```{r}

par (mfrow=c(1,2))
barplot(counts, main="Countries",col = rainbow(20), ylim=c(0,60), ylab="Diff of Monthly Avg. Temps.", las=2)
barplot(counts2, main="Cities",col = rainbow(20), ylim=c(0,60), ylab="Diff of Monthly Avg. Temps.", las=2)

#this is an interesting comment
```

***  
### __... Question 05__  
##### ... Semester End Bonus  
##### ... Run the following code in R and paste the final graph.  


```{r q5, echo = TRUE}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	t<-seq(0,10,length=1000)
	x<-sqrt(t)*cos(2*pi*t)
	y<-sqrt(t)*sin(2*pi*t)
	
	plot(x,y,axes=F,type="l",lwd=3,xlab="x(t)",ylab="y(t)",col="red")
	axis(1,at=seq(-3,3,by=0.5),labels=seq(-3,3,by=0.5))
	axis(2)
	
	box()
	
	title(main=expression(
	paste("(x(t),y(t)) with polar coordinates",
	(list(sqrt(t),2*pi*t))
	)))
```