---
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)

home_dir <- "~/_smu/_src/case_study_2"

data_dir <- paste0(home_dir, "/data")
setwd(home_dir)

```


***  
### __Question 02__  
Please watch videos1 and 2 in week 11 lecture assignment. You can download the code which
used for S&P from files tab.  
Please do the following with your assigned stock.  
- Download the data.  
- Calculate log returns.  
- Calculate volatility measure.  
- Calculate volatility over entire length of series for various three different decay factors.  
- Plot the results, overlaying the volatility curves on the data, just as was done in the S&P
example.  

 Group           | Stock 
---------------- | ------------- 
Patrick & Sunna  | GWPH        

```{r}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

library(tseries)

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ..	 volatility function
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

getVol <- function(d, logrets)
{
	var = 0
	lam = 0
	varlist <- c()

	for (r in logrets)
	{
		lam = lam*(1 - 1/d) + 1
		var = (1 - 1/lam)*var + (1/lam)*r^2
		varlist <- c(varlist, var)
	}
	sqrt(varlist)
}


# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	SNPdata <- get.hist.quote('gwph', quote = "Close")
	SNPret <- log(lag(SNPdata)) - log(SNPdata)
	SNPvol <- sd(SNPret) * sqrt(250) * 100


# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	recreate figure 6.12 in the text on page 155
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	volest  <- getVol(10, SNPret)
	volest2 <- getVol(30, SNPret)
	volest3 <- getVol(100, SNPret)
	
	plot(SNPdata)
	plot(log(SNPdata))
	plot(SNPret)

	plot(volest, type = "l")
	lines(volest2, type = "l", col = "red")
	lines(volest3, type = "l", col = "blue")


```

***  
### __Question 03__  

The built-in data set called Orange in R is about the growth of orange trees. The Orange data
frame has 3 columns of records of the growth of orange trees.  
Variable description  
* Tree : an ordered factor indicating the tree on which the measurement is made. The ordering
is according to increasing maximum diameter.  
* age : a numeric vector giving the age of the tree (days since 1968/12/31)  
* circumference : a numeric vector of trunk circumferences (mm). This is probably
“circumference at breast height”, a standard measurement in forestry.  
	
Submit your final R code and necessary plots for each part.


```{r}

data (Orange)

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	a) Calculate the mean and the median of the trunk circumferences
# ...		for different size of the trees. (Tree)  
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

col_drops <- c("age", "tree")
df_orange_tmp <- Orange[ , !(names(Orange) %in% col_drops)]
aggregate(df_orange_tmp[, 2], list(df_orange_tmp$Tree), mean)
aggregate(df_orange_tmp[, 2], list(df_orange_tmp$Tree), median)

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...b) Make a scatter plot of the trunk circumferences against the age of the tree.
# ...	Use different plotting symbols for different size of trees.  
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	plot(circumference ~ age, data = Orange,
		xlab = "Tree age (days since 1968/12/31)",
		ylab = "Tree circumference (mm)", las = 1,
		main = "Orange tree data",
		col = Tree)

	p <- ggplot(Orange, aes(age, circumference))
	p + geom_point(aes(color = factor(Tree)))

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	c) Display the trunk circumferences on a comparative boxplot against tree.
# ...	Be sure you order the boxplots in the increasing order of maximum diameter.  
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

	Orange$tree <- as.numeric(Orange$Tree)

	p <- ggplot(Orange, aes(x = as.numeric(Tree), y = circumference))
	p + geom_boxplot(aes(color = factor(tree)))

```

***  
